---
title: "UN Report on the relationship between GDP and Life Expectancy"
author: "Dongyuan Zhou"
date: "2023-10-13"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r package, include = FALSE}
library(tidyverse)
library(here)
library(knitr)

options(scipen = 100000)
```

# Summary

This report was prepared to the attention of the UN. It analyzes the relationship between a country's GDP, its life expectancy, and CO2 emissions. Our goal is to determine to what degree a country's economic strength or weakness is related to its public health status and impact on climate pollution. We hypothesize that both life expectancy and CO2 emissions increase with the country's GDP. 

## Important figure

```{r data, include = FALSE}
gapminder_1997 <- read_csv(here("gapminder_1997.csv"))
```
```{r gdp_lifeexp_1997, echo = FALSE}
ggplot(data = gapminder_1997, aes(x = gdpPercap, y = lifeExp, color = continent, size = pop/1000000))+
  geom_point()+
  labs(x = "GDP per Capita", 
       y = "Life Expectancy", 
       title = "Do people in wealthy countries live longer?",
       size = "Population (In Millions)")
```

```{r include = FALSE}
n_countries <- gapminder_1997 %>% 
  select(country) %>% 
  n_distinct()

minimum_gdp <- gapminder_1997 %>% 
  summarize(min(gdpPercap)) %>% 
  round(digits = 1)

maximum_gdp <- gapminder_1997 %>% 
  summarize(max(gdpPercap)) %>% 
  round(digits = 1)

# format - will format things nicely
```

The above plot shows the relationship between GDP per capita and life expectancy for a total of `r n_countries` countries. For this set of countries, economic wealth ranged from a minimum of \$`r minimum_gdp` to a maximum of \$`r maximum_gdp` per capita. 

## Related tables

| Summary of data | 
|:---------|----------:|
|Number of countries    | `r n_countries`|
|Minimum GDP per capita | `r minimum_gdp`|
|Maximum GDP per capita | `r maximum_gdp`|

```{r table, echo = FALSE}
gapminder_1997 %>% 
  filter(continent == "Oceania") %>% 
  kable()
```

## My number list
1. action 1 **R**
1. action 2 *gourp_by*
1. action 3 ***still R***
1. Markdown [cheatsheet](https://www.markdownguide.org/cheat-sheet/)

